Ordinal Models

### 2024-12-18 misc/ordinal.qmd
update - marginal effects

---
title: "Ordinal Models"
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  hugo-md:
    include: true
    html-math-method: mathjax
    output-file: ordinal_hugo.md
  gfm:
    echo: true
    output-file: ordinal.md
---

@@

pacman::p_load(dplyr,purrr,tidyr,here,tibble,stringr,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable,pander)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())
options(brms.backend="cmdstanr",mc.cores=4)

s1 <- readRDS(here::here("data/s1_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds"))))
s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))
```

@@

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                close_match = abs_error <= 0.03) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

@@

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id, state,refClass,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
                error = pct_change - pct,
                abs_error = abs(error)) |> 
      ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

ggplot(data = s2_agg1, aes(x = accuracy_level, fill = refClass)) +
  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Accuracy Levels by Reference Class",
       x = "Accuracy Level",
       y = "Percentage of Participants",
       fill = "Reference Class") +
  theme_minimal()

observed_props2 <- s2_agg1 |>
  group_by(refClass, accuracy_level) |>
  summarise(n = n()) |>
  group_by(refClass) |>
  mutate(prop = n/sum(n)) |>
  pivot_wider(
    names_from = accuracy_level,
    values_from = c(n, prop)
  )

  # Calculate proportions
prop_acc_s2 <- s2_agg1 %>%
    group_by(refClass, accuracy_level) %>%
    summarise(count = n()) %>%
    group_by(refClass) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

@@

ggplot(data = s2_agg1, aes(x = accuracy_level, fill = refClass)) +
  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Accuracy Levels by Reference Class",
       x = "Accuracy Level",
       y = "Percentage of Participants",
       fill = "Reference Class") +
  theme_minimal() + facet_wrap(~rounded+pct_goal)

prop4_acc_s2 <- s2_agg1 %>%
    group_by(refClass, accuracy_level,rounded,pct_goal) %>%
    summarise(count = n()) %>%
    group_by(refClass) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

@@

s2_agg1 %>%
    group_by( accuracy_level,rounded) %>%
    summarise(count = n()) %>%
    group_by(rounded) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

@@

s2_agg1 <- s2_agg1 %>%
  mutate(
    refClass = factor(refClass, levels = c("kWh", "Percentage", "USD")),
    rounded = factor(rounded, levels = c("Not Rounded", "Rounded")),
    pct_goal = factor(pct_goal, levels = c("10%", "15%"))
  )

ordinal_model_s2_probit <- brm(
  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("probit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_op.rds')) # Cache for efficiency
)

@@

pred_summary_s2 <- ordinal_model_s2_probit %>%
  epred_draws(newdata = s2_agg1, re_formula = NA) %>%
  group_by(refClass, rounded, pct_goal, .category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

@@

@@

# Fit the full model with interactions
ordinal_model_s2 <- brm(
  accuracy_level ~ refClass * rounded * pct_goal + (1|id) + (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = paste0(here::here("data/model_cache","s2_ordinal_full6.rds"))
)

pred_summary_s2 <- ordinal_model_s2 %>%
  epred_draws(newdata = s2_agg1, re_formula = NA) %>%
  group_by(refClass, rounded, pct_goal, .category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

@@

s2_agg1 <- s2_agg1 %>%
  mutate(
    refClass = factor(refClass, levels = c("kWh", "Percentage", "USD")),
    rounded = factor(rounded, levels = c("Not Rounded", "Rounded")),
    pct_goal = factor(pct_goal, levels = c("10%", "15%"))
  )

# Fit the full model with interactions
ordinal_model_s2 <- brm(
  accuracy_level ~ refClass * rounded * pct_goal + (1|id) + (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = paste0(here::here("data/model_cache","s2_ordinal_full6.rds"))
)

pred_summary_s2 <- ordinal_model_s2 %>%
  epred_draws(newdata = s2_agg1, re_formula = NA) %>%
  group_by(refClass, rounded, pct_goal, .category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

@@

>>>>

---
title: "Ordinal Models"
execute:
  #echo: false
  eval: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  hugo-md:
    include: true
    html-math-method: mathjax
    output-file: ordinal_hugo.md
  gfm:
    echo: true
    output-file: ordinal.md
---

@@

pacman::p_load(dplyr,purrr,tidyr,here,tibble,stringr,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable,pander,marginaleffects,ggstance)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())
options(brms.backend="cmdstanr",mc.cores=4)

# s1 <- readRDS(here::here("data/s1_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds"))))
# s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))

s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
  filter(!(id %in% readRDS(here::here("data/s1_grp_outlier_ids.rds")))) |>
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
  filter(!(id %in% readRDS(here::here("data/s2_grp_outlier_ids.rds"))) ) |>
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

@@

```{r}
s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.05 ~ "0.01-5% error",
                    TRUE ~ "Over 5% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-5% error", "Over 5% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.05 ~ "0.01-5% error",
                TRUE ~ "Over 5% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", ".01-5% error", "Over 5% error"),
            ordered = TRUE
        ))

@@

s2_agg <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(
    matched_goal = (pct_change == pct),
    close_match = abs(pct_change - pct) <= 0.02,
                error = pct_change - pct,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007)) |> 
      ungroup() |> 
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.05 ~ "0.01-5% error",
                    TRUE ~ "Over 5% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match","0.01-5% error", "Over 5% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.05 ~ ".01-5% error",
                TRUE ~ "Over 5% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "01-5% error", "Over 5% error"),
            ordered = TRUE
        ))

ggplot(data = s2_agg, aes(x = accuracy_level, fill = refClass)) +
  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Accuracy Levels by Reference Class",
       x = "Accuracy Level",
       y = "Percentage of Participants",
       fill = "Reference Class") +
  theme_minimal()

observed_props2 <- s2_agg |>
  group_by(refClass, accuracy_level) |>
  summarise(n = n()) |>
  group_by(refClass) |>
  mutate(prop = n/sum(n)) |>
  pivot_wider(
    names_from = accuracy_level,
    values_from = c(n, prop)
  )

  # Calculate proportions
prop_acc_s2 <- s2_agg %>%
    group_by(refClass, accuracy_level) %>%
    summarise(count = n()) %>%
    group_by(refClass) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

@@

ggplot(data = s2_agg, aes(x = accuracy_level, fill = refClass)) +
  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Accuracy Levels by Reference Class",
       x = "Accuracy Level",
       y = "Percentage of Participants",
       fill = "Reference Class") +
  theme_minimal() + facet_wrap(~rounded+pct_goal)

prop4_acc_s2 <- s2_agg %>%
    group_by(refClass, accuracy_level,rounded,pct_goal) %>%
    summarise(count = n()) %>%
    group_by(refClass) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

@@

s2_agg %>%
    group_by( accuracy_level,rounded) %>%
    summarise(count = n()) %>%
    group_by(rounded) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

@@

s2_agg <- s2_agg %>%
  mutate(
    refClass = factor(refClass, levels = c("kWh", "Percentage", "USD")),
    rounded = factor(rounded, levels = c("Not Rounded", "Rounded")),
    pct_goal = factor(pct_goal, levels = c("10%", "15%"))
  )

ordinal_model_s2_probit <- brm(
  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
  data = s2_agg,
  family = cumulative("probit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_op.rds')) # Cache for efficiency
)

@@

pred_summary_s2 <- ordinal_model_s2_probit %>%
  epred_draws(newdata = s2_agg, re_formula = NA) %>%
  group_by(refClass, rounded, pct_goal, .category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

@@

marginal_effects(ordinal_model_s2_probit)

## S2 logit version
```{r}

ordinal_model_s2_logit <- brm(
  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
  data = s2_agg,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_add.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit)
# Regression Coefficients:
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept[1]          -2.13      0.64    -3.39    -0.86 1.00      639     1328
# Intercept[2]          -0.62      0.63    -1.89     0.63 1.00      658     1366
# Intercept[3]           3.15      0.65     1.88     4.42 1.00      717     1638
# refClassPercentage     0.83      0.75    -0.64     2.33 1.00      452     1253
# refClassUSD            1.87      0.79     0.31     3.35 1.00      508     1260
# roundedRounded        -0.66      0.18    -1.01    -0.31 1.00     4276     3095
# pct_goal15%           -0.44      0.18    -0.79    -0.10 1.00     6263     3109

pred_summary_s2 <- ordinal_model_s2_logit %>%
  epred_draws(newdata = s2_agg, re_formula = NA) %>%
  group_by(refClass, rounded, pct_goal, .category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

### logit vs. probit

loo_logit <- loo(ordinal_model_s2_logit)
loo_probit <- loo(ordinal_model_s2_probit)
loo_compare(loo_logit, loo_probit)

waic(ordinal_model_s2_logit)
waic(ordinal_model_s2_probit)

pp_check(ordinal_model_s2_logit, type = "bars_grouped", group="refClass", fatten = 2) +
  scale_x_continuous("Response Category", breaks = 1:4, 
            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  ggtitle("Posterior Predictive Check by Reference Class - logit") +
  theme_minimal() +
  theme(
    legend.background = element_blank(),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1))

pp_check(ordinal_model_s2_probit, type = "bars_grouped", group="refClass", fatten = 2) +
  scale_x_continuous("Response Category", breaks = 1:4, 
            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  ggtitle("Posterior Predictive Check by Reference Class - probit") +
  theme_minimal() +
  theme(
    legend.background = element_blank(),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1))

@@

# Fit the full model with interactions
ordinal_model_s2 <- brm(
  accuracy_level ~ refClass * rounded * pct_goal + (1|id) + (1|state),
  data = s2_agg,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = paste0(here::here("data/model_cache","s2_ordinal_full6.rds"))
)

pred_summary_s2 <- ordinal_model_s2 %>%
  epred_draws(newdata = s2_agg, re_formula = NA) %>%
  group_by(refClass, rounded, pct_goal, .category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

@@

s2_agg <- s2_agg %>%
  mutate(
    refClass = factor(refClass, levels = c("kWh", "Percentage", "USD")),
    rounded = factor(rounded, levels = c("Not Rounded", "Rounded")),
    pct_goal = factor(pct_goal, levels = c("10%", "15%"))
  )

# Fit the full model with interactions
ordinal_model_s2 <- brm(
  accuracy_level ~ refClass * rounded * pct_goal + (1|id) + (1|state),
  data = s2_agg,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = paste0(here::here("data/model_cache","s2_ordinal_full6.rds"))
)

pred_summary_s2 <- ordinal_model_s2 %>%
  epred_draws(newdata = s2_agg, re_formula = NA) %>%
  group_by(refClass, rounded, pct_goal, .category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

@@

# 3 level

```{r}
s1_aggb <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.05 ~ "0.01-5% error",
                    TRUE ~ "Over 5% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-5% error", "Over 5% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

ggplot(data = s1_aggb, aes(x = refClass, fill = accuracy_level)) +
  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Accuracy Levels by Reference Class",
       x = "Accuracy Level",
       y = "Percentage of Participants",
       fill = "Reference Class") +
  theme_minimal()

library(lmerTest)
mixed_model_abs_error <- lmer(log_abs_error ~ refClass+calc+ (1+plan|id) + (1|state), data = s1_aggb)
summary(mixed_model_abs_error)

logistic_model_clean_mixed <- glmer(matched_goal ~ refClass +calc+ (1|id) + (1|state), data = s1_aggb, family = binomial)
summary(logistic_model_clean_mixed)
emmeans::emmeans(logistic_model_clean_mixed, pairwise ~ refClass)

mixed_model_abs_error <- lmer(log_abs_error ~  refClass+calc +(1|state)+ (1|id), data = s1_aggb)
summary(mixed_model_abs_error)
emmeans::emmeans(mixed_model_abs_error, pairwise ~ refClass)

# categorial mixed effects regression on accuracy_level (ordered factor) with clean data (frequentist)
ordinal_model_clean_mixed <- clmm(accuracy_level ~ refClass +calc + (1|id) + (1|state), data = s1_aggb)
summary(ordinal_model_clean_mixed)
emmeans::emmeans(ordinal_model_clean_mixed, pairwise ~ refClass)

ordinal_model_s1 <- brm(
    accuracy_level ~ refClass +calc + (1|id) + (1|state),
    data = s1_aggb,
    family = cumulative("logit"),
    cores = 4,
    iter = 4000,
    control = list(adapt_delta = 0.98), 
    prior = c(prior(normal(0, 4), class = "Intercept"), 
                prior(normal(0, 4), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_acc3_add.rds')) 
)
# summary(ordinal_model_s1)
#  Family: cumulative 
#   Links: mu = logit; disc = identity 
# Formula: accuracy_level ~ refClass + calc + (1 | id) + (1 | state) 
#    Data: s1_aggb (Number of observations: 940) 
#   Draws: 4 chains, each with iter = 4000; warmup = 2000; thin = 1;
#          total post-warmup draws = 8000
# 
# Multilevel Hyperparameters:
# ~id (Number of levels: 235) 
#               Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# sd(Intercept)     3.90      0.35     3.27     4.64 1.00     2072     4098
# 
# ~state (Number of levels: 4) 
#               Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# sd(Intercept)     0.58      0.45     0.12     1.78 1.00     2961     2957
# 
# Regression Coefficients:
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept[1]          -4.21      0.85    -5.90    -2.58 1.00     1567     2898
# Intercept[2]          -0.89      0.82    -2.49     0.71 1.00     1528     2564
# refClassPercentage     1.44      0.71     0.07     2.88 1.00     1348     2624
# refClassUSD            3.13      0.69     1.81     4.50 1.00     1373     2564
# calcUsedCalculator    -3.30      0.73    -4.80    -1.92 1.00     1491     2765

pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2)
pp_check(ordinal_model_s1, type = "dens_overlay_grouped", group="refClass", fatten = 2,ndraws=100)

ame1 <- avg_slopes(
  ordinal_model_s1, 
  variables = "refClass"
)
print(ame1)
 #         Group         Contrast Estimate   2.5 %   97.5 %
 # Exact match   Percentage - kWh  -0.1064 -0.2164 -0.00504
 # Exact match   USD - kWh         -0.2001 -0.2988 -0.10898
 # 0.01-5% error Percentage - kWh  -0.0211 -0.0577  0.00242
 # 0.01-5% error USD - kWh         -0.0811 -0.1143 -0.04856
 # Over 5% error Percentage - kWh   0.1314  0.0063  0.24907
 # Over 5% error USD - kWh          0.2842  0.1723  0.38086

as.data.frame(fixef(ordinal_model_s1)[,-2])|> as.data.frame() %>%
    rownames_to_column(var = "Parameter") %>%
    mutate(across(where(is.numeric), exp)) |>
    filter(!stringr::str_detect(Parameter, "Intercept")) |> 
    filter(!stringr::str_detect(Parameter, "calc")) |> 
    # rename columns to |comparison           | odds_ratio| ci_lower| ci_upper|
    rename(Comparison = Parameter, odds_ratio = Estimate, ci_lower = Q2.5, ci_upper = Q97.5) |>
    # rename levels to Percentage vs kWh; USD vs kWh; 
    mutate(Comparison = case_when(
        str_detect(Comparison, "refClassPercentage") ~ "Percentage vs kWh",
        str_detect(Comparison, "refClassUSD") ~ "USD vs kWh",
        TRUE ~ Comparison
    )) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"))

# |Comparison        |odds_ratio |ci_lower |ci_upper |
# |:-----------------|:----------|:--------|:--------|
# |Percentage vs kWh |4.2        |1.1      |18       |
# |USD vs kWh        |22.9       |6.1      |90       |
```

condEffects <- function(m,xvar){
  m |> ggplot(aes(x = {{xvar}}, y = .value, color = refClass, fill = refClass)) + 
  stat_dist_pointinterval() + 
  stat_halfeye(alpha=.1, height=.5) +
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

}
```

s2_agg <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(
    matched_goal = (pct_change == pct),
    close_match = abs(pct_change - pct) <= 0.02,
                error = pct_change - pct,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007)) |> 
      ungroup() |> 
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.05 ~ "0.01-5% error",
                    TRUE ~ "Over 5% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match","0.01-5% error", "Over 5% error"),
                ordered = TRUE
            )
        )

ggplot(data = s2_agg, aes(x = refClass, fill = accuracy_level)) +
  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Accuracy Levels by Reference Class",
       x = "Accuracy Level",
       y = "Percentage of Participants",
       fill = "Reference Class") +
  theme_minimal()

library(lmerTest)
mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s2_agg)
summary(mixed_model_abs_error)

logistic_model_clean_mixed <- glmer(matched_goal ~ refClass+calc+pct_goal+rounded +(1|state)+ (1|id) , data = s2_agg, family = binomial)
summary(logistic_model_clean_mixed)
emmeans::emmeans(logistic_model_clean_mixed, pairwise ~ refClass)

mixed_model_abs_error <- lmer(log_abs_error ~  refClass+calc+pct_goal+rounded +(1|state)+ (1|id), data = s2_agg)
summary(mixed_model_abs_error)
emmeans::emmeans(mixed_model_abs_error, pairwise ~ refClass)

library(ordinal)    
library(parameters)

# categorial mixed effects regression on accuracy_level (ordered factor) with clean data (frequentist)
ordinal_model_clean_mixed <- clmm(accuracy_level ~ refClass + calc+pct_goal+rounded + (1|id) + (1|state), data = s2_agg)
summary(ordinal_model_clean_mixed)
emmeans::emmeans(ordinal_model_clean_mixed, pairwise ~ refClass)

ordinal_model_clean_mixed <- clmm(accuracy_level ~ (refClass*pct_goal*rounded)+ calc + (1|id) + (1|state), data = s2_agg)
summary(ordinal_model_clean_mixed)
emmeans::emmeans(ordinal_model_clean_mixed, pairwise ~ refClass)
pairs(emmeans(ordinal_model_clean_mixed, ~ refClass | rounded +pct_goal+calc))

ordinal_model_clean_mixed <- clmm(accuracy_level ~ refClass + rounded + pct_goal + calc + refClass:calc+refClass:rounded + refClass:pct_goal + rounded:pct_goal + (1|id) + (1|state), data = s2_agg)
summary(ordinal_model_clean_mixed)
emmeans::emmeans(ordinal_model_clean_mixed, pairwise ~ refClass+calc)

emms_ref <- emmeans(ordinal_model_clean_mixed, ~ refClass, type = "link")
contrast_ref <- contrast(emms_ref, "trt.vs.ctrl", ref = "kWh") 
odds_ratios <- summary(contrast_ref, type = "response") # back-transform to odds
odds_ratios

pairs(emmeans(ordinal_model_clean_mixed, ~ refClass | rounded +pct_goal+calc))
pairs(emmeans(ordinal_model_clean_mixed, ~ refClass+calc))

# ordinal regression
polr_model <- MASS::polr(accuracy_level ~ refClass + calc, data = s2_agg)
summary(polr_model)
emmeans::emmeans(polr_model, pairwise ~ refClass+calc)

m_equal <- clm(accuracy_level ~ refClass+calc+pct_goal+rounded, 
    data = s2_agg, 
    link="probit")

parameters::model_parameters(m_equal) |> 
  insight::print_html()
summary(m_equal)

ame_by_rounded_pct_goal <- avg_slopes(
  m_equal,
  variables = "refClass",
  by = c("rounded", "pct_goal", "accuracy_level")
)
